---
title: "Lokaverkefni nafn hér"
author: "Bergrós Skúladóttir, Indriði Arnaldsson & Íris Aníta Eyþórsdóttir"
output:
  prettydoc::html_pretty:
    theme : hpstr
    highlight: github
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, error = F)
```

```{css, echo=FALSE, class.output="scroll-100"}
.scroll-100 {
  max-height: 300px;
  overflow-y: auto;
  bakground-color:inherit;
}
pre[class] {
  max-height: 160px;
}
```

```{r}
library(tidyverse)
library(ggridges)
library(data.table)
library(gridExtra)
library(corrplot)
library(kableExtra)
library(glmnet)
options(scipen = 999)
library(randomForest)
library(gbm)
library(scales)
```

## Introduction 
In this assignment we will carry out an analysis of a dataset containing information of about 40 thousand real estates sold in Iceland from 2008 to 2018. The origin of the data can be traced to the National Registry of Iceland which performs an evaluation of housing prices every year. 

In the first part of the assignment we will explore the data visually. In the second part we will predict sale prices using x and x methods. In the third part we will predict the values of x (categorical variable) using the x and x methods. 

The dataset contains many variables, some of which have been excluded from analysis. For this assignment we will focus on data concerning the Capital Region of Iceland. The resulting dataset contains approximately 35 thousand observations and 17 variables regarding housing prices, municipality, age of construction, property type and property area among other. We have included only properties that are single-family houses or apartment buildings. 

## Reading data and cleanup
```{r}
data <- read_delim("https://www.skra.is/library/Samnyttar-skrar-/Fyrirtaeki-stofnanir/Fasteignamat-2019/gagnasafn_ib_2018.csv", ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"),
                                     trim_ws = TRUE)
data <- data %>% select(kdagur, nuvirdi, byggar, lyfta, ibm2, fjhaed, fjbilsk, fjbkar, fjsturt, fjklos, fjeld, fjherb, fjstof, fjgeym, matssvaedi, ibteg, svfn) %>%
  filter(matssvaedi < 1000) %>%
  mutate(lyfta = ifelse(lyfta >= 1, 1, 0)) %>%
  mutate(ibteg = ifelse(ibteg == 11, 1, 0)) %>%
  mutate_at("lyfta", factor) %>%
  mutate_at("matssvaedi", as.character) %>%
  mutate_at("ibteg", factor) 
         
  
data$svfn <- recode_factor(data$svfn, `0000` = "Reykjavik", `1000`= "Kopavogur", `1100` = "Seltjarnarnes",`1300` = "Gardabaer", `1400` = "Hafnarfjordur", `1604` = "Mosfellsbaer")
data$kdagur <- as.Date(data$kdagur, format = "%d.%m.%Y")
data$yearkeypt <- format(data$kdagur, format="%Y")
data$yearkeypt <- as.numeric(data$yearkeypt)
data$age <- (as.numeric(data$yearkeypt)-data$byggar)
data <- select(data, -kdagur, -byggar)
data$matssvaedi <- recode_factor(data$matssvaedi, `11`="Rvk:Vesturbær", `20` = "Rvk:Miðbær", `25`="Rvk:Miðbær", `31`="Rvk:Miðbær", `70`="Rvk:Vesturbær", `72`="Rvk:Vesturbær", `75`="Rvk:Vesturbær", `80`="Rvk:Hlíðar", `85`="Rvk:Bústaðarhverfi", `90`="Rvk:Laugardalur", `91`="Rvk:Bústaðarhverfi", `100`="Rvk:Laugardalur", `110`="Rvk:Grafarvogur", `120`="Rvk:Grafarvogur", `130`="Rvk:Grafarvogur", `140`="Rvk:Grafarvogur", `150`="Rvk:Breiðholt", `160`="Rvk:Breiðholt", `161`="Rvk:Breiðholt", `170`="Rvk:Breiðholt", `180`="Rvk:Grafarholt", `181`="Rvk:Grafarholt", `200`="Rvk:Árbær", `210`="Rvk:Árbær", `220`="Rvk:Árbær", `270`="Rvk:Árbær", `280`="Rvk:Bústaðarhverfi", `281`="Rvk:Bústaðarhverfi", `282` ="Rvk:Bústaðarhverfi", `283`="Rvk:Bústaðarhverfi", `284`="Rvk:Bústaðarhverfi", `290`="Rvk:Kjalarnes",
                                 `300`="Kóp:Vesturbær",`320`="Kóp:Austurbær",`330`="Kóp:Smárar", `340`="Kóp:Lindir", `350`="Kóp:Hvörf",`351`="Kóp:Kórar",
                                 `400`="Seltjarnarnes", `500`="Grb:Miðbær", `510`="Grb:Ásar", `511`="Grb:Arnarnes", `520`="Grb:Sjáland", `530`="Grb:Hraunholt", `540`="Grb:Arnarnes", `550`="Grb:Urriðarholt", `560`="Grb:Flatir", `590`="Grb:Annað",`700`="Grb:Álftanes", `600`="Hfj:Miðbær", `620`="Hfj:Holt", `630`="Hfj:Vellir", `640`="Hfj:Ásland", `650`="Hfj:Setberg", `660`="Hfj:Flensborg", `670`="Hfj:Vellir", `680`="Hfj:Holt", `800`="Mos:Miðbær", `810`="Mos:Teigar", `820`="Mos:Leirvogstunga", `830`="Mos:Dalur", `840`="Mos:Hlíðar", `850`="Mos:Helgafell", `890`="Mos:Annað", `999`="Kóp:Annað")
```

## Part I: Exploratory Analysis

In this part we will explore the data. Property price is an important variable in our dataset and we are interested in exploring the relationship it has with the other variables in our data set. 

```{r, fig.align='center'}
#color theme
ColorsDT <-  data.table(Group=LETTERS[1:50], Color=c("#e0bba8","#d6a78e","#cd9274", "#c37e59","#b86a41","#9d5b38","#834c2f","#693c25","#4e2d1c","#341e12","#efc699","#eab57a", "#e5a45b","#e1923c","#d88121","#b96e1c","#9a5c17","#7b4912","#5c370e","#3d2409","#d6bdb2", "#caa99a","#bd9583","#b1816c","#a36e56","#8c5e4a","#744e3d","#5d3f31","#462f25","#2e1f18", "#e8bda0","#e1a983","#da9566","#d48149","#ca6e2f","#ad5e28","#904e22","#733f1b","#562f14", "#391f0d","#d9bfae","#ceab96","#c3987d","#b88465","#aa714e","#926143","#7a5138","#61412d", "#493021","#302016"), key="Group")
set.seed(4)
#núvirði skoðað útfrá:
#(1)svfn
f1 <- data %>% ggplot(aes(x=nuvirdi/1e3,y=svfn, fill=svfn)) +
  geom_density_ridges_gradient() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
        legend.position = "none") + 
  scale_x_continuous(labels = label_number(suffix = " m.ISK"), limits=c(0,100), breaks=seq(0,100,by=50)) +
  scale_fill_manual(values=ColorsDT$Color) + 
  labs(y = " ", x = "Property Price")
#(2)matssvaedi
f2 <- data %>% ggplot(aes(x=nuvirdi/1e3,y=matssvaedi, fill=svfn)) +
  geom_density_ridges_gradient() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
        legend.position = "none") + 
  ggtitle(" ") +   scale_x_continuous(labels = label_number(suffix = " m.ISK"), limits=c(0,100), breaks=seq(0,100,by=50)) +
  scale_fill_manual(values=ColorsDT$Color) + 
  labs(y = " ", x = "Property Price")

grid.arrange(f1,f2,nrow=1,top="Figure 1: Property price based on municipality and neighbourhood")
```

In figure 1 we see the property prices based on the municipality on the one hand, and neighbourhood on the other. Overall, the property prices seem to have a fairly normal distribution across the municipalities. Property prices in Reykjavík appear to be the lowest, perhaps because the properties are smaller. The prices for properties in two neighbourhoods in Garðabær and in Hlíðar appear to be, on average, higher than other neighbourhoods in our dataset. 

```{r, fig.align = 'center'}
#(3)yearkeypt
data %>% 
  ggplot(aes(x=nuvirdi/1e3,y=as.factor(yearkeypt), fill=as.factor(yearkeypt))) +
  geom_density_ridges_gradient() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
        legend.position = "none", plot.title = element_text(hjust = 0.5)) + 
  ggtitle(" ") + scale_x_continuous(labels = label_number(suffix = " m.ISK"), limits=c(0,100), breaks=seq(0,100,by=20)) + 
  scale_fill_manual(values=ColorsDT$Color) + 
  labs(y = " ", x = "Property Price") + 
  ggtitle("Figure 2: Property price by year of sale")
```

In figure 2 we can see the property price by year of sale. Property prices have gradually increased over the years. The average property price over the years 2012 to 2018 is approximately 37 million isk ($\mu \approx$ `r round(mean(data$nuvirdi))`).  

```{r, fig.align = 'center'}
#(4)fj breytunar
ff <- function(x){
data %>%
  group_by(bla = as.factor(x)) %>%
  summarise(price = round(mean(nuvirdi/1e3))) %>% 
   mutate(Label=paste0(price), "m.ISK") %>%
  ggplot(aes(x=price,y=bla, fill=bla)) + 
    geom_col() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), 
          legend.position = "none") +
    scale_fill_manual(values=sample(ColorsDT$Color)) + geom_text(aes(label=price), nudge_x = -10, col = "white")
}
p <- scale_x_continuous(labels = label_number(suffix = " m.ISK"), limits=c(0,200))
q <- scale_x_continuous(labels = label_number(suffix = " m.ISK"), limits=c(0,140), breaks = seq(0,140,by=65))
grid.arrange(ff(data$fjhaed)+ 
               labs(y = "Floors", x = " ")+p, ff(data$fjbilsk) + 
               labs(y = "Parking spaces", x = " ")+p, ff(data$fjbkar) + 
               labs(y = "Baths", x = " ")+p, ff(data$fjsturt) + 
               labs(y = "Showers", x = " ")+p, ff(data$fjklos) + 
               labs(y = "Toilets", x = " ")+p, ff(data$fjeld) + 
               labs(y = "Kitchens", x = " ")+p, top= "Figure 3: Mean property price by various elements")
grid.arrange(ff(data$fjstof) + 
               labs(y = "Living Rooms", x = " ")+q, ff(data$fjgeym) + 
               labs(y = "Storage Rooms", x = " ")+q, ff(data$fjherb) + 
               labs(y = "Rooms", x = " ")+q, nrow=1)
```

Here above we see plots of property price according to various property attributes. For example, properties that have 5 toilets are more expensive than those with fewer toilets, and properties with 3 kitchens appear to be the most expensive. 

```{r, fig.align = 'center'}
#(5)íbuðir með/án lyftu (bara fjölbýlishús)
# búum til nýtt gagnasafn sem inniheldur bara fjölbýlishús
data_fjolbylishus <- data %>% filter(ibteg == 0) %>%
  mutate(lyfta = ifelse(lyfta == 1, "Yes", "No")) %>%
  group_by(lyfta,svfn) %>%
  summarise(nuvirdi=mean(nuvirdi))

data_fjolbylishus %>%
  ggplot() + geom_bar(aes(y=nuvirdi/1e3,x=svfn, fill=lyfta), stat= "identity", position = "dodge") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) + 
  ggtitle("Figure X: Mean prices for apartments by municipality and elevator status") + 
  scale_y_continuous(labels = label_number(suffix = " m.ISK"), limits=c(0,50), breaks=seq(0,50,by=10)) + 
  scale_fill_manual(name = "Elevator", values=sample(ColorsDT$Color)) + 
  labs(x = " ", y = "Property Price")

datt <- data_fjolbylishus %>%
  pivot_wider(names_from = svfn, values_from=nuvirdi)

dat <- datt[-1]
Heild <- apply(dat,2,sum)
mism <- c(dat[2,])-(dat[1,])
tt <- paste(round((mism)/Heild,2)*100,"%")
kable(t(tt), col.names = c("RVK","KÓP","SEL","GRB","HFJ","MOS"), caption = "Table 1: Percentage increase in apartment price by having an elevator") %>% kable_styling(full_width = F)
```

In figure 3 we see the average property price of apartments based on whether the building has an elevator or not. Apartments in buildings with an elevator are on average more expensive. In table 1 we see the percentage increase in apartments that are in buildings with an elevator. Seltjarnarnes is the only municipality where having an elevator does not increase the propery price. This could peh  

```{r, fig.align = 'center'}
# verð út frá aldri byggingar og tegund húsnæðis.
f3 <- data %>% 
  mutate(ibteg=ifelse(data$ibteg == 1, "Single-family house", "Apartment building"))
ggplot(f3, aes(x=nuvirdi/1e3,y=age, group = ibteg, color = ibteg)) +
  geom_smooth() + 
  ggtitle("Figure 4: Property price based on age and type of property") +
  labs(x = "Property price", y = "Age of property") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), 
        plot.title = element_text(hjust = 0.5)) +
   scale_x_continuous(labels = label_number(suffix = " m.ISK"), limits=c(0,100), breaks=seq(0,100,by=25)) + 
    scale_color_manual(name = "Type of property",values = sample(ColorsDT$Color))
```

In figure 4 we have the average property price based on age and type of property. Single-family houses are on average more expensive than apartment buildings and the older the property the more inexpensive it is. Age appears to be irrelevant for single-family houses priced 50 million isk and above.   

```{r, fig.align = 'center'}
# price per square meter by municipality
data <- data %>% 
  mutate(fermetraverd = nuvirdi/ibm2)

ggplot(data, aes(x = fermetraverd, y = svfn, fill = svfn)) +
  geom_density_ridges_gradient() +
  labs(x= "Price per square meter", y = "Municipality") +
  ggtitle("Figure 5: Price per square meter by municipality") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), legend.position = "none") + 
  scale_fill_manual(values = sample(ColorsDT$Color)) + 
  scale_color_manual(name = "Type of property",values=sample(ColorsDT$Color)) +
  xlim(c(0, 75))
```

In figure 5 we see price per square meter by municipality. We can gather that prices per square meter are a bit higher for properties in Garðabær and Seltjarnarnes (and Mosfellsbær to a degree) than for properties in the other municipalities. 

```{r}
#Corrplot allar numeric breytur
da <- data %>%
  dplyr::select(where(is.numeric))
da <- na.omit(da)
colnames(da) <- c("Pr.Price","Sqm","Floors","Parking","Baths","Showers","Toilets","Kitchens","Rooms","Liv.r.", "Storage","Sale.Year", "Age", "Sqm.Price")
par(xpd=TRUE)
corrplot.mixed(cor(da), lower.col = "#d88121", number.cex=0.7, tl.pos = "lt", upper.col = "#8c5e4a", tl.col="#341e12", tl.cex=0.7, tl.srt = 45, title = "Figure 6: Correlation plot of property attributes", mar=c(0,0,4,0))
```

A correlation plot of 12 property attributes can be seen on figure 6. Property price appears to have a moderately strong positive correlation with area (property size in square meters) and number of rooms, living rooms, floors, showers and toilets in a property. 

The last five attributes mentioned here above also appear to have a moderately strong postive relationship with each other. 

## Part II: Predicting sale prices 

In this part we will predict sales prices using the Lasso and boosting methods. 

#### The lasso

```{r}
set.seed(4)
blabb <- subset(data[-13])

x <- model.matrix(nuvirdi~., blabb)[,-1]
y <- as.matrix(blabb$nuvirdi)
grid <- 10^seq(2,-4, by=-0.1)
train <- sample(1:nrow(x),nrow(x)*0.7)
Tr <- blabb[train,]
y <- Tr$nuvirdi

lasso.cv <- cv.glmnet(x[train,],
                      y,
                      alpha=1,
                      lambda = grid)

bestlam <- lasso.cv$lambda.min
data.frame(lambda = lasso.cv$lambda, 
           cv.mse = lasso.cv$cvm, 
           nonzero = lasso.cv$nzero) %>%
  mutate(minmse = as.numeric(lambda == lasso.cv$lambda.min)) %>%
  ggplot(aes(x = lambda, y = cv.mse,
             col = ifelse(lambda==minmse, "#493021","#d88121"))) + 
  geom_point(size=3) + 
  geom_line() +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
               labels = trans_format("log10", math_format(10^.x))) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
               labels = trans_format("log10", math_format(10^.x))) +
  scale_color_identity() + xlab( expression(paste("log(", lambda, ")"))) + ylab("Cross-Validation MSE") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), legend.position = "none") + ggtitle("Figure X: Lasso Cross-Validation") + annotate(geom="text",y=10^8.14379, x=10, label= expression(paste(lambda, "=log(10)")))
bestlam
```

Our optimal value of $\lambda$ is log(`r bestlam`) for minimum MSE.

```{r}
library(reshape2)
lasso.mod <- glmnet(x[train,],
                    y,
                    alpha=1,
                    standardize = T,
                    metric="mse",
                    lambda = bestlam)
beta <- coef(lasso.mod)
plot <- as.data.frame(as.matrix(beta))
plot$coef <- row.names(plot)
plot <- reshape2::melt(plot, id="coef")
plot$variable <- as.numeric(gsub("s","",plot$variable))
plot$lambda <- lasso.mod$lambda[plot$variable+1]
haha <- plot[-1,] %>%
  filter(value>0) %>%
  arrange(-value)
rownames(haha) <- c("Seltjarnarnes", "Garðabær", "Parking", "Toilets", "Prop. Type", "Sales Year", "Showers", "Living rooms", "Elevator", "Rooms", "Storage", "Bathtubs", "Kópavogur", "Price per Sqm", "Sqm")

haha %>%
  ggplot(aes(value,reorder(rownames(haha),value), col=coef)) + geom_point(show.legend = F)+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), legend.position = "none") + 
  scale_fill_manual(values = sample(ColorsDT$Color)) + 
  scale_color_manual(name = "Type of property",values=sample(ColorsDT$Color)) + labs(x="Coefficient", y=" ") + ggtitle(expression(paste("Figure X: Predictors with non-zero Lasso coefficients with ", lambda, " = log(10)")))

lasso.pred <- predict(lasso.mod, s=bestlam, newx=x[train,])
lasso.trainmse <- round(mean((lasso.pred[1]-data$nuvirdi))/1e3,2)
```

The Lasso approach found 15 important coefficients. They are estimated to explain around `r percent(lasso.mod$dev)` of variance in property price. With $\lambda$ set to 10. With training error of around `r lasso.trainmse`m.ISK. 

  The main issue with this approach is blindness regarding correlated/confounding variables.  
  
```{r}
x.test <- x[-train,]
lasso.pred <- predict(lasso.mod, s=bestlam, newx=x.test)
lassomse <- round(mean(y-lasso.pred[1]),2)

gaga <- cbind(lasso.pred,y)
gaga <- as.data.frame(gaga)
colnames(gaga) <- c("pred", "true")
gaga %>%
  ggplot(aes(x=pred/1e3,y=true/1e3)) + geom_point()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), legend.position = "none") +  labs(y="Observed", x="Predicted") + ggtitle("Figure X: Lasso predicted versus observed property prices") + geom_abline(slope=1,intercept = 0,lty=2, col= "#d88121") +
scale_x_continuous(labels = label_number(suffix = " m.ISK"), limits=c(0,150)) +
  scale_y_continuous(labels = label_number(suffix = " m.ISK"), limits=c(0,150))
```

The Lasso test error obtained was `r round(lassomse/1e3,2)`m.ISK. 


#### Boosting

Fitting a boosted model to our training data. Tuning the lambda parameter to the lowest test MSE.
```{r, warning = F, message = F, echo = F}
# train og x.test
set.seed(1)

data.Log.nuvirdi <- data
data.Log.nuvirdi$nuvirdi <- log(data.Log.nuvirdi$nuvirdi) # nota logarithmann af 

TRAIN <- sample(1:nrow(data.Log.nuvirdi), nrow(data.Log.nuvirdi)*0.7)
train.set <- data.Log.nuvirdi[TRAIN,]
test.set <- data.Log.nuvirdi[-TRAIN,]

# Tuning lambda
lambdaGrid <- seq(-3, -0.9, length.out = 20) # Typical values are 0.01 or 0.001
lambdas <- 10^lambdaGrid

test.MSE <- rep(NA, length(lambdas)) #define list to store training errors

for (i in 1:length(lambdas)) {
  GBM.boost <- gbm(nuvirdi ~ ., data = train.set,
                     distribution = "gaussian", 
                     shrinkage = lambdas[i])
 
   boost.predict <- predict(GBM.boost, test.set, 
                                    n.trees = 1000)
   test.MSE[i] <- round(mean((boost.predict - test.set$nuvirdi)^2), 2)
}

plot(lambdas, test.MSE, 
     type = "b", 
     xlab = "Shrinkage value (lambda)", 
     ylab = "Test MSE", 
     main = "Figure X: Test MSE for a range of lambda values", 
     col = "#d88121")

min.testMSE <- min(test.MSE) # gives the lowest test MSE
min.lambda <- round(lambdas[which.min(test.MSE)], 2) # gives the lambda corresponding to lowest test
```

Our optimal $\lambda$ is `r min.lambda` for the minimum test MSE. The minimum test MSE is `r min.testMSE`. 

Finna út prediction og reikna út prediction accuracy
```{r}
boost <- gbm(nuvirdi ~ ., data = train.set, #fjarlægja fermetraverð breytuna því hún er búin til úr núvirði og ibm2
             distribution = "gaussian", 
             shrinkage = 0.1, # common lambda to use (0.01 or 0.001)
             n.trees = 1000)
library(vip)
vip(boost, aesthetics = list(color = "#d88121", fill = "#a36e56"), mapping = aes_string(fill = "Variable")) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),
        plot.title = element_text(hjust = 0.5)) + 
  ggtitle("Figure X: Relative influence of variables using boosting") + 
  labs(x = "Property Price", y = "Relative influence") 

#  scale_y_discrete(labels=c("fjhaed"="Hæðir")) Virkar ekki. 


#summary(boost, cBars = 10, 
#        method = relative.influence, 
#        las = 2) #breyta þessu í ggplot. 


#Reikna út hversu % af predictive power - er það venjulega gert í boosting? 

```

LAGA: liti og y-axis á plottinu. 

There are three variables that have high relative importance, they are square meters of property, the year the property was bought and number of bathrooms. Square meters per property is by far the most important variable. In figure X and X in part I we learned that the more square meters and bathrooms are in a property the more expensive it is. We also learned that property prices increase the younger the property is.

Fjalla hér aðeins um muninn á test MSE hjá Lasso og Boosting. Bera saman hvaða breytur voru taldar mikilvægastar þegar kemur að því að spá núvirði. 


