---
title: "Lokaverkefni nafn hér"
author: "Bergrós Skúladóttir, Indriði Arnaldsson & Íris Aníta Eyþórsdóttir"
output:
  prettydoc::html_pretty:
    theme : hpstr
    highlight: github
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, error = F)
```

```{css, echo=FALSE, class.output="scroll-100"}
.scroll-100 {
  max-height: 300px;
  overflow-y: auto;
  bakground-color:inherit;
}
pre[class] {
  max-height: 160px;
}
```

```{r}
library(tidyverse)
library(ggridges)
library(data.table)
library(gridExtra)
library(corrplot)
library(kableExtra)
library(glmnet)
options(scipen = 999)
library(randomForest)
library(gbm)
```

## Introduction 
In this assignment we will carry out an analysis of a dataset containing information of about 40 thousand real estates sold in Iceland from 2008 to 2018. The origin of the data can be traced to the National Registry of Iceland which performs an evaluation of housing prices every year. 

In the first part of the assignment we will explore the data visually. In the second part we will predict sale prices using x and x methods. In the third part we will predict the values of x (categorical variable) using the x and x methods. 

The dataset contains many variables, some of which have been excluded from analysis. For this assignment we will focus on data concerning the Capital Region of Iceland. The resulting dataset contains approximately 35 thousand observations and 17 variables regarding housing prices, municipality, age of construction, property type and property area among other. We have included only properties that are single-family houses or apartment buildings. 

## Reading data and cleanup
```{r}
data <- read_delim("https://www.skra.is/library/Samnyttar-skrar-/Fyrirtaeki-stofnanir/Fasteignamat-2019/gagnasafn_ib_2018.csv", ";", escape_double = FALSE, locale = locale(encoding = "ISO-8859-1"),
                                     trim_ws = TRUE)
data <- data %>% select(kdagur, nuvirdi, byggar, lyfta, ibm2, fjhaed, fjbilsk, fjbkar, fjsturt, fjklos, fjeld, fjherb, fjstof, fjgeym, matssvaedi, ibteg, svfn) %>%
  filter(matssvaedi < 1000) %>%
  mutate(lyfta = ifelse(lyfta >= 1, 1, 0)) %>%
  mutate(ibteg = ifelse(ibteg == 11, 1, 0)) %>%
  mutate_at("lyfta", factor) %>%
  mutate_at("matssvaedi", as.character) %>%
  mutate_at("ibteg", factor) 
         
  
data$svfn <- recode_factor(data$svfn, `0000` = "Reykjavik", `1000`= "Kopavogur", `1100` = "Seltjarnarnes",`1300` = "Gardabaer", `1400` = "Hafnarfjordur", `1604` = "Mosfellsbaer")
data$kdagur <- as.Date(data$kdagur, format = "%d.%m.%Y")
data$yearkeypt <- format(data$kdagur, format="%Y")
data$yearkeypt <- as.numeric(data$yearkeypt)
data$age <- (as.numeric(data$yearkeypt)-data$byggar)
#number of negative values for property age - properties bought unfinished 
nrow(data[data$age<0,])
data <- select(data, -kdagur, -byggar)
data$matssvaedi <- recode_factor(data$matssvaedi, `11`="Rvk:Vesturbær", `20` = "Rvk:Miðbær", `25`="Rvk:Miðbær", `31`="Rvk:Miðbær", `70`="Rvk:Vesturbær", `72`="Rvk:Vesturbær", `75`="Rvk:Vesturbær", `80`="Rvk:Hlíðar", `85`="Rvk:Bústaðarhverfi", `90`="Rvk:Laugardalur", `91`="Rvk:Bústaðarhverfi", `100`="Rvk:Laugardalur", `110`="Rvk:Grafarvogur", `120`="Rvk:Grafarvogur", `130`="Rvk:Grafarvogur", `140`="Rvk:Grafarvogur", `150`="Rvk:Breiðholt", `160`="Rvk:Breiðholt", `161`="Rvk:Breiðholt", `170`="Rvk:Breiðholt", `180`="Rvk:Grafarholt", `181`="Rvk:Grafarholt", `200`="Rvk:Árbær", `210`="Rvk:Árbær", `220`="Rvk:Árbær", `270`="Rvk:Árbær", `280`="Rvk:Bústaðarhverfi", `281`="Rvk:Bústaðarhverfi", `282` ="Rvk:Bústaðarhverfi", `283`="Rvk:Bústaðarhverfi", `284`="Rvk:Bústaðarhverfi", `290`="Rvk:Kjalarnes",
                                 `300`="Kóp:Vesturbær",`320`="Kóp:Austurbær",`330`="Kóp:Smárar", `340`="Kóp:Lindir", `350`="Kóp:Hvörf",`351`="Kóp:Kórar",
                                 `400`="Seltjarnarnes", `500`="Grb:Miðbær", `510`="Grb:Ásar", `511`="Grb:Arnarnes", `520`="Grb:Sjáland", `530`="Grb:Hraunholt", `540`="Grb:Arnarnes", `550`="Grb:Urriðarholt", `560`="Grb:Flatir", `590`="Grb:Annað",`700`="Grb:Álftanes", `600`="Hfj:Miðbær", `620`="Hfj:Holt", `630`="Hfj:Vellir", `640`="Hfj:Ásland", `650`="Hfj:Setberg", `660`="Hfj:Flensborg", `670`="Hfj:Vellir", `680`="Hfj:Holt", `800`="Mos:Miðbær", `810`="Mos:Teigar", `820`="Mos:Leirvogstunga", `830`="Mos:Dalur", `840`="Mos:Hlíðar", `850`="Mos:Helgafell", `890`="Mos:Annað", `999`="Kóp:Annað")
```

## Exploratory Analysis

In this part we will explore the data. Property price is an important variable in our dataset and we are interested in exploring the relationship it has with the other variables in our data set. 

```{r, fig.align='center'}
#color theme
ColorsDT <-  data.table(Group=LETTERS[1:50], Color=c("#e0bba8","#d6a78e","#cd9274", "#c37e59","#b86a41","#9d5b38","#834c2f","#693c25","#4e2d1c","#341e12","#efc699","#eab57a", "#e5a45b","#e1923c","#d88121","#b96e1c","#9a5c17","#7b4912","#5c370e","#3d2409","#d6bdb2", "#caa99a","#bd9583","#b1816c","#a36e56","#8c5e4a","#744e3d","#5d3f31","#462f25","#2e1f18", "#e8bda0","#e1a983","#da9566","#d48149","#ca6e2f","#ad5e28","#904e22","#733f1b","#562f14", "#391f0d","#d9bfae","#ceab96","#c3987d","#b88465","#aa714e","#926143","#7a5138","#61412d", "#493021","#302016"), key="Group")
set.seed(4)
#núvirði skoðað útfrá:
#(1)svfn
f1 <- data %>% ggplot(aes(x=nuvirdi,y=svfn, fill=svfn)) +
  geom_density_ridges_gradient() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.position = "none") + 
  xlim(c(0,100000)) + 
  scale_fill_manual(values=ColorsDT$Color) + 
  labs(y = " ", x = "Property Price")
#(2)matssvaedi
f2 <- data %>% ggplot(aes(x=nuvirdi,y=matssvaedi, fill=svfn)) +
  geom_density_ridges_gradient() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.position = "none") + 
  ggtitle(" ") + xlim(c(0,100000)) + scale_fill_manual(values=ColorsDT$Color) + 
  labs(y = " ", x = "Property Price")
grid.arrange(f1,f2,nrow=1,top="Figure 1: Property price based on municipality and neighbourhood")
```

In figure 1 we see the property prices based on the municipality on the one hand, and neighbourhood on the other. Overall, the property prices seem to have a fairly normal distribution across the municipalities. Property prices in Reykjavík appear to be the lowest, perhaps because the properties are smaller. The prices for properties in two neighbourhoods in Garðabær and in Hlíðar appear to be, on average, higher than other neighbourhoods in our dataset. 

```{r, fig.align = 'center'}
#(3)yearkeypt
data %>% ggplot(aes(x=nuvirdi,y=as.factor(yearkeypt), fill=as.factor(yearkeypt))) +
  geom_density_ridges_gradient() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), legend.position = "none") + 
  ggtitle(" ") + xlim(c(0,100000)) + scale_fill_manual(values=ColorsDT$Color) + 
  labs(y = " ", x = "Property Price") + ggtitle("Figure 2: Property price by year of sale")
```
ATH: tók burt aðra myndina svo þarf að laga textann i samræmi við það!

In figure 2 we can see the property price by year of sale. The vertical line represents the mean property price over the years 2012 to 2018 (Overall $\mu \approx$ `r round(mean(data$nuvirdi))`). Property prices have gradually increased over the years.  

```{r, fig.align = 'center'}
#(4)fj breytunar
ff <- function(x){
data %>%
  group_by(bla = as.factor(x)) %>%
  summarise(price = mean(nuvirdi)) %>% 
   mutate(Label=paste0(round(price/10000,2), "m")) %>%
  ggplot(aes(y=price,x=bla, fill=bla)) + 
    geom_col() +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), 
          legend.position = "none") + ylim(c(0,100000)) + 
    scale_fill_manual(values=sample(ColorsDT$Color)) + geom_text(aes(label=Label), nudge_y = 5)
    }
grid.arrange(ff(data$fjhaed)+ 
               labs(x = "Floors", y = " "), ff(data$fjbilsk) + 
               labs(x = "Parking spaces", y = " "), ff(data$fjbkar) + 
               labs(x = "Baths", y = " "), ff(data$fjsturt) + 
               labs(x = "Showers", y = " "), ff(data$fjklos) + 
               labs(x = "Toilets", y = " "), ff(data$fjeld) + 
               labs(x = "Kitchens", y = " "), ff(data$fjherb) + 
               labs(x = "Rooms", y = " "), ff(data$fjstof) + 
               labs(x = "Living Rooms", y = " "), ff(data$fjgeym) + 
               labs(x = "Storage Rooms", y = " "), top="Figure 3: Mean property price by various elements")
```
LAGA - uppsetninguna i grid.arrange og labels.. 

Here above we see plots of property price according to various property attributes. For example, properties that have 5 toilets are more expensive than those with fewer toilets, and properties with 3 kitchens appear to be the most expensive. 

```{r, fig.align = 'center'}
#(5)íbuðir með/án lyftu (bara fjölbýlishús)
# búum til nýtt gagnasafn sem inniheldur bara fjölbýlishús
data_fjolbylishus <- data %>% filter(ibteg == 0) %>%
  mutate(lyfta = ifelse(lyfta == 1, "Yes", "No")) %>%
  group_by(lyfta,svfn) %>%
  summarise(nuvirdi=mean(nuvirdi))
data_fjolbylishus %>%
  ggplot() + geom_bar(aes(x=nuvirdi,y=svfn, fill=lyfta), stat= "identity", position = "dodge") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) + 
  ggtitle("Figure 3: Mean prices for apartments by municipality and elevator status") + 
  xlim(c(0, 45000)) + 
  scale_fill_manual(name = "Elevator", values=sample(ColorsDT$Color)) + 
  labs(y = " ", x = "Property Price") + coord_flip()

datt <- data_fjolbylishus %>%
  pivot_wider(names_from = svfn, values_from=nuvirdi)
dat <- datt[-1]
Heild <- apply(dat,2,sum)
mism <- c(dat[2,])-(dat[1,])
tt <- paste(round((mism)/Heild,2)*100,"%")
kable(t(tt), col.names = c("RVK","KÓP","SEL","GRB","HFJ","MOS"), caption = "Table 1: Percentage increase in apartment price by having an elevator") %>% kable_styling(full_width = F)
```

LAGA - setti þetta bara upp í töflu því mér datt í hug að gætum gert eitthvað sniðugt við þessar upplýsingar en veit ekki nákvæmlega hvað (setja inn sem label a myndina? eða tala um í texta bara..)

In figure 3 we see the average property price of apartment buildings based on whether the building has an elevator or not. Apartment buildings with elevators are on average more expensive.  

```{r, fig.align = 'center'}
# verð út frá aldri byggingar og tegund húsnæðis.
f3 <- data %>% 
  mutate(ibteg=ifelse(data$ibteg == 1, "Single-family house", "Apartment building"))
ggplot(f3, aes(x=nuvirdi,y=age, group = ibteg, color = ibteg)) +
  geom_smooth() + 
  ggtitle("Figure 4: Property price based on age and type of property") +
  labs(x = "Property price", y = "Age of property") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
   xlim(c(0,100000)) + 
    scale_color_manual(name = "Type of property",values = sample(ColorsDT$Color))
```

In figure 4 we have the average property price based on age and type of property. Single-family houses are on average more expensive than apartment buildings and the older the property the more inexpensive it is. Age appears to be irrelevant for single-family houses priced 50 million isk and above.   

```{r, fig.align = 'center'}
# price per square meter by municipality
data <- data %>% 
  mutate(fermetraverd = nuvirdi/ibm2)
ggplot(data, aes(x = fermetraverd, y = svfn, fill = svfn)) +
  geom_density_ridges_gradient() +
  labs(x= "Price per square meter", y = "Municipality") +
  ggtitle("Figure 5: Price per square meter by municipality") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), legend.position = "none") + 
  scale_fill_manual(values = sample(ColorsDT$Color)) + 
  scale_color_manual(name = "Type of property",values=sample(ColorsDT$Color)) +
  xlim(c(0, 75))
```

In figure 5 we see price per square meter by municipality. We can gather that prices per square meter are a bit higher for properties in Garðabær and Seltjarnarnes (and Mosfellsbær to a degree) than for properties in the other municipalities. 

```{r}
#Corrplot allar numeric breytur
da <- data %>%
  dplyr::select(where(is.numeric))
head(da)
colnames(da) <- c("Pr.Price","Sqm","Floors","Parking","Baths","Showers","Toilets","Kitchens","Rooms","Liv.r.", "Storage","Salesyear","Age")
corrplot.mixed(cor(da[-13]), lower.col = "#d88121", number.cex=0.7, tl.pos = "lt", upper.col = "#8c5e4a", tl.col="#341e12", tl.cex=0.7, tl.srt = 45, 
               title = "Figure 6: Correlation plot of property attributes", mar=c(0,0,4,0))
```

LAGA - label cutoffs 

A correlation plot of 12 property attributes can be seen on figure 6. Property price appears to have a moderately strong positive correlation with area (property size in square meters) and number of rooms, living rooms, floors, showers and toilets in a property. 

The last five attributes mentioned here above also appear to have a moderately strong postive relationship with each other. 

## Part II: Predicting sale prices 

In this part we will predict sales prices using the Lasso and boosting methods. 

#### The lasso

ÞURFUM AÐ BREYTA AÐEINS TRAIN OG TEST SET ÞANNIG AÐ ÞAÐ SÉ HÆGT AÐ NOTA FYRIR ALLAR AÐFERÐIR. ÞAÐ ER SNIÐIÐ AÐ LASSO AÐFERÐINNI EINS OG ER, GET EKKI NOTAÐ SAMA TRAIN OG TEST SET Í T.D. BOOSTING. Breytan x (fyrir model.matrix skipunina er einnig óvenju stór. Þurfum að endurhugsa aðeins hvernig við viljum kóða fyrir lasso aðferðina)

Bjuggum til ný test og training set (sjá undir Boosting kaflanum). Ef við höldum okkur við það þá ættum við að geta notað sama test og training set fyrir allar aðferðir. 
```{r}
set.seed(4)
x <- model.matrix(nuvirdi~., data)[,-1]
y <- log(data$nuvirdi)
grid <- 10^seq(2,-2, by= -.1)
train <- sample(1:nrow(x),nrow(x)*0.7)

lasso.cv <- cv.glmnet(x[train,],
                      y[train],
                      alpha=1,
                      standardize=T,
                      lambda = grid)

bestlam <- lasso.cv$lambda.min 

data.frame(lambda = lasso.cv$lambda, 
           cv.mse = lasso.cv$cvm, 
           nonzero = lasso.cv$nzero) %>%
  ggplot(aes(x = lambda, y = cv.mse, col = ifelse(lambda==bestlam, "#d88121", "#c3987d"))) + 
  geom_point() + 
  geom_line() +
  scale_x_continuous(trans = 'log10') + 
  theme(legend.position = "bottom") + 
  scale_color_identity() +
  labs(x = "Lambda", 
       y = "Cross-Validation MSE", 
       col = "Non-Zero Coefficients:", 
       title = "Figure 7: 10-Fold CV Lasso lambda selection")+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), legend.position = "none")

bestlam
```
HUGMYND, bæta bestlam á plottið. 
Our optimal value of $\lambda$ is `r bestlam` for minimum MSE.

```{r}
library(reshape2)
lasso.mod <- glmnet(x[train,],
                    y[train],
                    alpha=1,
                    standardize = T,
                    lambda = bestlam)
beta <- coef(lasso.mod)
plot <- as.data.frame(as.matrix(beta))
plot$coef <- row.names(plot)
plot <- reshape2::melt(plot, id="coef")
plot$variable <- as.numeric(gsub("s","",plot$variable))
plot$lambda <- lasso.mod$lambda[plot$variable+1]
plot[-1,] %>%
  filter(value>0) %>%
  ggplot(aes(value,reorder(coef,value), col=coef)) + geom_point(show.legend = F)+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), legend.position = "none") + 
  scale_fill_manual(values = sample(ColorsDT$Color)) + 
  scale_color_manual(name = "Type of property",values=sample(ColorsDT$Color)) + labs(x="Coefficient", y=" ") + ggtitle("Figure X: Predictors with non-zero Lasso coefficients")
  
mean((plot$value-data$nuvirdi)^2)
```

```{r}
library(scales)
x.test <- x[-train,]
lasso.pred <- predict(lasso.mod, s=bestlam, newx=x.test)
lassomse <- percent(mean(lasso.pred[1]-y)^2)

gaga <- cbind(lasso.pred,y)
gaga <- as.data.frame(gaga)
colnames(gaga) <- c("pred", "true")
gaga %>%
  ggplot(aes(y=exp(pred)/1000,x=exp(true)/1000)) + geom_point()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_blank(), legend.position = "none") +  labs(x="Observed", y="Predicted") + ggtitle("Figure X: Lasso predicted versus observed property prices") + geom_abline(slope=1,intercept = 0,lty=2, col= "#d88121")
```

LAGA: bæta við "m.kr" á axis values

The Lasso test error obtained was `r lassomse`. 

#### Boosting

```{r, warning = F, message = F, echo = F}
# train og x.test
set.seed(1)
 
TRAIN <- sample(1:nrow(data), nrow(data)*0.7)
train.set <- data[TRAIN,]
test.set <- data[-TRAIN,]


lambdaGrid <- seq(-3, -2, length.out = 20) # Typical values are 0.01 or 0.001
lambdas <- 10^lambdaGrid

test.MSE <- rep(NA, length(lambdas)) #define list to store training errors

for (i in 1:length(lambdas)) {
  GBM.boost <- gbm(nuvirdi ~ ., data = train.set,
                     distribution = "gaussian", 
                     n.trees = 1000, 
                     shrinkage = lambdas[i])
 
   boost.predict <- predict(GBM.boost, test.set, 
                                    n.trees = 1000)
   test.MSE[i] <- round(mean((boost.predict - test.set$nuvirdi)^2), 2)
}

plot(lambdas, test.MSE, 
     type = "b", 
     xlab = "Shrinkage value (lambda)", 
     ylab = "Test MSE", 
     main = "Test MSE for a range of lambda values", 
     col = "darkblue")
```
ATH, keyrslutími á knitr er mjög hægur við að keyra boosting. Ef við vitum að kóðinn virkar getum við tekið þessa skipun út í bili svo knitting taki ekki of langan tíma. 
